---
title: "Data Collection"
output: 
  html_document:
    includes:
      after_body: ../docs/footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting data for the forecasting

In this section we're `collecting the data` for the target variable (the variable 
we want to forecast) and for potentially useful explanatory variables. 
Since the goal is to forecast, we will have to have least 250 observations 
at our disposal. Based on the frequency of observation time horizon will be 
different.

That means we could use daily, weekly, monthly (we'll need at least 21 years) 
or quarterly data (we'll need at least 63 years).

We settled on using daily data from period of `01-01-2019` to `06-10-2021`, with 
`616 observations` in total, nearly two and a half years of trading data.

For the analysis, our `target variable` is `Microsoft (MSFT) stock`. Some of the 
`explanatory variables` (used for regression modeling and forecasting and for
initial data analysis) are:

- S&P500
- Nasdaq
- AAPL
- GOOG
- IBM
- 3M

For data collection process we're using 
[quantmod library](https://cran.r-project.org/web/packages/quantmod/quantmod.pdf).

```{r echo=FALSE, warning=FALSE}
# Importing libraries
suppressMessages(library(quantmod))
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
options("getSymbols.warning4.0"=FALSE)
```



```{r echo=FALSE, warning=FALSE}
# Date period for historical prices
start_date = "2019-01-01"
end_date = "2021-06-11"
```



```{r echo=FALSE, warning=FALSE}
get_and_process_ticker_data <- function(ticker_, from_, to_, output_path_){
  
  # Getting ticker data for selected date range 
  data <- getSymbols(ticker_, from = from_, to= to_, auto.assign = FALSE, return.class = "data.frame")
  
  # Printing data set
  print(tail(data))
  
  # Add auxiliary column for plotting
  data$date <- as.Date(rownames(data))
  
  # Save to file
  write.csv(x=data, file=output_path_)
  
  ticker_ <- gsub("\\^", "", ticker_)
  
  # Plotting data set
  ggplot(data=data, aes_string(x="date", y=paste0(ticker_, ".Close"))) + 
        geom_line(color="blue") +
        ggtitle(paste(ticker_, "historical price")) + 
        xlab("Date") + 
        ylab("Price") +
        theme(plot.title = element_text(hjust = 0.5))
  
}
```


### Microsoft data
```{r echo=FALSE, warning=FALSE}
get_and_process_ticker_data("MSFT", start_date, end_date, "../data/MSFT.csv")
```

### S&P500 data
```{r echo=FALSE, warning=FALSE}
get_and_process_ticker_data("^GSPC", start_date, end_date, "../data/S&P500.csv")
```


### Nasdaq data
```{r echo=FALSE, warning=FALSE}
get_and_process_ticker_data("^IXIC", start_date, end_date, "../data/Nasdaq.csv")
```

### Apple data
```{r echo=FALSE, warning=FALSE}
get_and_process_ticker_data("AAPL", start_date, end_date, "../data/AAPL.csv")
```

### Google data
```{r echo=FALSE, warning=FALSE}
get_and_process_ticker_data("GOOG", start_date, end_date, "../data/GOOG.csv")
```

### IBM data
```{r echo=FALSE, warning=FALSE}
get_and_process_ticker_data("IBM", start_date, end_date, "../data/IBM.csv")
```

### 3M data
```{r echo=FALSE, warning=FALSE}
get_and_process_ticker_data("MMM", start_date, end_date, "../data/MMM.csv")
```

