---
title: "StockAnalysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

## MSFT stock analysys

In this section we will do analysis of `MSF returns`, perform various statistical
tests in order to better understand characteristics of MSFT returns and its 
distribution during last two and a half years `01/01/2019 - 06/11/2021`.

```{r}
# Importing libraries
suppressMessages(library(quantmod))
suppressMessages(library(fBasics))
```


### Dataset

In the table we can see with which data we're working on. Data sets used in this
analysis are obtained with `quantmod` library.
```{r}
# Reading data set
MSFT <- read.csv(file = "../data/MSFT.csv", row.names = 1, header = TRUE)

# Printing data
tail(MSFT[,1:5])
```


### Visualizing the dataset

Let's now plot the stock price for that period. 
As you can see, only by `visually inspecting plots` to have a better feeling of 
the MSFT price movement, without performing any statistical tests, it seems that
`before H1 of 2020` (when Covid-19 starting to take its toll), there was a clear
`upwards trend`. An as mentioned, because of the global pandemic there was a
`sudden drop in price`. Later on we will compare MSFT performance compared to the
S&P500 and Nasdaq indexes and some of MSFT's competitors in the tech sector.
What is also interesting is that `in the H2 of 2020 and in the H1 of 2021`, 
we can clearly see again the `upwards trend`, which resulted in `stock price surge`.

On the second plot, we inspect both the `volume` and the `stock price`. First thing
that's obvious is that there was a major `sell-off in the H1 of 2020` when Covid
pandemic stuck. What is also interesting is that is we closely examine the plot
we will see that occasionally there are huge volumes during some trading sessions
that moved the price, regardless of the direction.
```{r}
# Converting price series to xts objects
# (in order to work with quantmod library)
MSFT_xts <- xts(MSFT[, 1:5], order.by=as.POSIXct(MSFT$date))

# Plotting only closing price
chartSeries(MSFT_xts[,4],theme="white",major.tick="months")

# Plotting prices and volume
chartSeries(MSFT_xts,theme="white",major.tick="months") 
```

## MSFT returns

In this section we'll work with MSFT returns, performing various statistical
tests to figure out and estimate the distribution of returns and thus to have 
a better idea of stocks risk-reward ratio, potential tail risks and other things.

### Calculating and visualizing the returns

Let's first calculate the simple returns for MSFT:

- Daily returns
- Weekly returns
- Monthly returns
- Quarterly returns
- Yearly returns
```{r}
# Calculating returns

MSFT_xts.retDaily <- periodReturn(MSFT_xts, period = "daily")
MSFT_xts.retWeekly <- periodReturn(MSFT_xts, period = "weekly")
MSFT_xts.retMonthly <- periodReturn(MSFT_xts, period = "monthly")
MSFT_xts.retQuarterly <- periodReturn(MSFT_xts, period = "quarterly")
MSFT_xts.retYearly <- periodReturn(MSFT_xts, period = "yearly")
```

Now let's plot previously calculated returns. There are some interesting things
we can see on the plots:

- Daily returns - we can notice some volatility clustering meaning that there
  are periods with high volatility, periods with low volatility and periods with
  moderate volatility. We automatically know that there isn't a constant variance
  of this returns, so there is some time variation in volatility.
- Weekly returns - we can see similar behaviour here as well, but it isn't so
  obvious as for the daily returns. For both returns we can clearly see that
  the period with the highest volatility was in H1 of the 2020 for obvious reasons.
  In that period there were both the highest and the lowest daily and weekly 
  returns for two and a half year period.
- Quarterly returns - as previously mentioned H1 of 2020 was the most volatile
  period and that was the time when the whole market went down and in the transitioning
  period of Q2-Q3 market started to recover. What is interesting is that Q1 of 
  that year had a lowest quarterly return (nearly 0%) and the Q2 of the same year
  had the highest quarterly return (nearly 30%) in the time horizon of two and a 
  half years, which represents our data set.
- Yearly returns - the most important thing to take it from here is that the 
  MSFT stock returns in 2019 (astonishing nearly 60%) outperformed both the returns
  in 2020 (around 40%) and in H1 of 2021 (around 5%). 
```{r}
# Plotting returns
chartSeries(MSFT_xts.retDaily, theme = "white", up.col="blue", name = "MSFT - Daily returns")
chartSeries(MSFT_xts.retWeekly, theme = "white", up.col="blue", name = "MSFT - Weekly returns")
chartSeries(MSFT_xts.retMonthly, theme = "white", up.col="blue", name = "MSFT - Montly returns")
chartSeries(MSFT_xts.retQuarterly, theme = "white", up.col="blue", name = "MSFT - Quarterly returns")
chartSeries(MSFT_xts.retYearly, theme = "white", up.col="blue", name = "MSFT Yearly returns")
```

### Analysing daily returns


Here we can see some more detailed statistical properties of our data set. Some
of the interesting thigs we could notice are:

- The highest (14.21%) and the lowest (-14.73%) return MSFT had in one day during
  period od last 615 days.
- Simple estimation of future daily return as the mean is 0.17%
- From the first quantile (-0.0072) and the third quantile (0.0111), which is a 
  magnitude larger, we have an indicator that returns could be positively skewed,
  more leaning towards positive returns, but we'll statistically calculate that
  in the one of the next steps.
```{r}
# Summary statistics
basicStats(MSFT_xts.retDaily)
```

Let's examine `average` and `annualized` daily return and volatility.
Expected annual return is `42.2%` while expected annual volatility is `32.4%`.
```{r}
# Volatility

avg_return <- mean(MSFT_xts.retDaily)
vol <- stdev(MSFT_xts.retDaily)

print(c("Average daily return:", round(avg_return, 4)))
print(c("Annualized return:", round(avg_return * 252, 3)))
print(c("Volatility:", round(vol, 3)))
print(c("Annualized volatility:", round(vol * sqrt(252), 3)))


```


We calculated that the expected daily return is `0.18%`. Let's do zero mean test
(null hypothesis is that returns are zero on average). We got that `p-value` for 
the test is `0.0335` which is lower than the `significance level of 0.5` for
rejecting the null hypothesis. Because of the we reject the null hypothesis and
accept the alternative.

Previously we already calculated average daily return, which is not zero. 
But because it's not so large number, we got the p-value for the test to be 0.033, 
which means that if the daily return was smaller than that, it wouldn't be statistically significant and we would probably accept the null hypothesis. In that case we 
could say with statistical certainty that average daily return is 0.
```{r}
# Testing zero mean return
t.test(MSFT_xts.retDaily) 
```


#### Central moments

Third central moment (`skewness`) is `-0.093`. As skewness is negative and small,
returns are slightly skewed to the left. We'll also perform skewness test.

Obtained p-value for the test is less than significance level so we reject the null
hypothesis (H0 is that it's normal distribution, thus skewness is 0).
```{r}
SK = skewness(MSFT_xts.retDaily)  # Skewness
T = length(MSFT_xts.retDaily)     # Sample size
t3 = SK/sqrt(6/T)                 # Skewness test
pval_S = 2*(1-pnorm(abs(t3)))     # p-value

print(c("Skewness:", SK))
print(c("P-value:", pval_S))
```


Fourth central moment (`kurtosis`) is `10.13`. In this case we're calculating
excess kurtosis kurtosis, the reference value is 0, the same as for the skewness.
As kurtosis is positive and large number, returns distribution is `leptokurtic`.
That means that distribution has `fatter tails`, indicating potential `tail risk`.
Returns that are more than three standard deviations is considered to instantiate
tail risk.

Similarly as for the skewness, we'll also perform kurtosis test. 
Obtained p-value for the test is less than significance level so we reject the null
hypothesis (H0 is that it's normal distribution, thus that kurtosis is 0).
```{r}
K = kurtosis(MSFT_xts.retDaily)  # Kurtosis
t4 = K/sqrt(24/T)                # Kurtosis test
pval_K = 2*(1-pnorm(abs(t4)))     # Compute p-value

print(c("Kurtosis:", K))
print(c("P-value:", pval_K))
```

#### Distribution of returns

For simplification purposes, especially in academia, it is often assumed that 
returns follow `normal distribution`. As we previously seen from calculating 
kurtosis, assuming normal distribution could be dangerous because it ignores 
`tail risks` for rare and extreme events.

`Jarque-Bera normality test` is statistical test which checks if distribution
follows normal distribution. For MSFT returns we got that `p-value` is really small
(`2.2e-16`), so we reject the null hypothesis. This test takes both deviation in
the skewness and in the kurtosis. 

```{r}
normalTest(MSFT_xts.retDaily,method='jb')
```

To have a better felling of all the metrics we calculated previously we can plot
the `histogram of MSFT daily returns`.

Just by looking at the histogram, if we remove outliers, we could say that returns
follow normal distribution. As skewness is nearly equal to zero, histogram looks
quite symmetric.

```{r}
hist(MSFT_xts.retDaily, nclass=35) 
```

In order to better compare distribution of MSFT daily returns to normal distribution
we will estimate PDF and plot it together with normal distribution.

What we previously calculated from Jarque-Bera normality test, we can clearly
see that MSFT returns are deviating from normal distribution.
```{r}
# Obtain density estimate
d1 = density(MSFT_xts.retDaily)

# Create a sequence of x with increment 0.001.
x = seq(-.4,.4,.001)

# Plot the density
plot(d1$x, d1$y, xlab='rtn', ylab='density', type='l') 

# Create artificial data from Normal distribution
# with mean and variance equal to sample values of MSFT returns
y1 = dnorm(x, mean(MSFT_xts.retDaily), stdev(MSFT_xts.retDaily)) 

# Add density of normal r.v.
lines(x, y1, lty=2)
```

