---
title: "StockAnalysis"
output: 
  html_document:
    includes:
      after_body: ../docs/footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

## MSFT stock analysys

In this section we will do analysis of `MSF returns` perform various statistical
tests in order to better understand characteristics of MSFT returns and its 
distribution during last two and a half years from `01/01/2019` to `06/11/2021`.

```{r echo=FALSE}
# Importing libraries
suppressMessages(library(quantmod))
suppressMessages(library(fBasics))
suppressMessages(library(np))
```


### Dataset

In the table we can see with which data we're working on. Data sets used in this
analysis are obtained with `quantmod` library.
```{r echo=FALSE}
# Reading data set
MSFT <- read.csv(file = "../data/MSFT.csv", row.names = 1, header = TRUE)

# Printing data
tail(MSFT[,1:5])
```


### Visualizing the dataset

Let's now plot the stock price for that period. 
As you can see, only by `visually inspecting plots`, in order to have a better feeling of 
the MSFT price movement, without performing any statistical tests, it seems that
`before H1 of 2020` (when Covid-19 starting to take its toll), there was a clear
`upwards trend`. And as mentioned, because of the global pandemic, there was a
`sudden drop in price`. Later on, we will compare MSFT performance to the S&P500 
indexes that we will use as a benchmark.

What is also interesting is that `in the H2 of 2020 and in the H1 of 2021`, 
we can again clearly see the `upwards trend`, which resulted in `stock price surge`.

On the second plot, we inspect both the `volume` and the `stock price`. First thing
that's obvious is that there was a major `sell-off in the H1 of 2020` when Covid
pandemic stuck. What is also interesting is that is we closely examine the plot
we will see that occasionally there are huge volumes during some trading sessions
that moved the price, regardless of the direction.
```{r echo=FALSE}
# Converting price series to xts objects
# (in order to work with quantmod library)
MSFT_xts <- xts(MSFT[, 1:5], order.by=as.POSIXct(MSFT$date))
```

```{r echo=FALSE}
# Plotting only closing price
chartSeries(MSFT_xts[,4], theme="white", up.col="blue", major.tick="months", name="MSFT - Closing price")

```

```{r echo=FALSE}
# Plotting prices and volume
chartSeries(MSFT_xts, theme="white", major.tick="months", name="MSFT - Prices and Volume") 
```


## MSFT returns

In this section we'll work with MSFT returns, performing various statistical
tests to figure out and estimate the distribution of returns, thus to have 
a better idea of stocks risk-reward ratio, potential tail risks and other 
interesting things.

### Calculating and visualizing the returns

Let's first calculate the simple returns for MSFT:

- Daily returns
- Weekly returns
- Monthly returns
- Quarterly returns
- Yearly returns
```{r echo=FALSE}
# Calculating returns

MSFT_xts.retDaily <- periodReturn(MSFT_xts, period = "daily")
MSFT_xts.retWeekly <- periodReturn(MSFT_xts, period = "weekly")
MSFT_xts.retMonthly <- periodReturn(MSFT_xts, period = "monthly")
MSFT_xts.retQuarterly <- periodReturn(MSFT_xts, period = "quarterly")
MSFT_xts.retYearly <- periodReturn(MSFT_xts, period = "yearly")
```

Now let's plot previously calculated returns. There are some interesting things
we can see on the plots:

- `Daily returns` - we can notice some volatility clustering, meaning that there
  are periods with high volatility, periods with low volatility and periods with
  moderate volatility. We automatically know that there isn't a constant variance
  of this returns, there is some time variation in volatility.
- `Weekly returns` - we can see similar behaviour here as well, but it isn't so
  obvious as for the daily returns. For both the daily and weekly returns we can 
  clearly see that the period with the highest volatility was in H1 of the 2020,
  for obvious reasons. In that period there were both the highest and the lowest
  daily and weekly returns, for two and a half year period.
- `Quarterly returns` - as previously mentioned H1 of 2020 was the most volatile
  period and that was the time when the whole market went down and in the transitioning
  period of Q2-Q3, market started to recover. What is interesting is that Q1 of 
  that year had a lowest quarterly return (nearly 0%) and the Q2 of the same year
  had the highest quarterly return (nearly 30%) in the time horizon of two and a 
  half years, which represents our data set.
- `Yearly returns` - the most important thing to take it from here is that the 
  MSFT stock returns in 2019 (astonishing nearly 60%) outperformed both the returns
  in 2020 (around 40%) and in H1 of 2021 (around 5%). 


Plotting returns:

```{r echo=FALSE}
chartSeries(MSFT_xts.retDaily, theme = "white", up.col="blue", name = "MSFT - Daily returns")
```

```{r echo=FALSE}
chartSeries(MSFT_xts.retWeekly, theme = "white", up.col="blue", name = "MSFT - Weekly returns")
```

```{r echo=FALSE}
chartSeries(MSFT_xts.retMonthly, theme = "white", up.col="blue", name = "MSFT - Montly returns")
```

```{r echo=FALSE}
chartSeries(MSFT_xts.retQuarterly, theme = "white", up.col="blue", name = "MSFT - Quarterly returns")
```

```{r echo=FALSE}
chartSeries(MSFT_xts.retYearly, theme = "white", up.col="blue", name = "MSFT Yearly returns")
```

### Analysing daily returns


Here we can see some statistical properties of our data set in more detail. Some
of the interesting things we could notice are:

- The highest daily return (`14.21%`) and the lowest daily (`-14.73%`) return 
  MSFT stock had during the period of 615 trading days.
- Simple estimation of future MSFT daily return, calculated as the mean, is `0.17%`
- From the first quantile (-0.0072) and the third quantile (0.0111), we have an 
  indicator that returns are somewhere symmetric, with small potential skewness. 
  But we'll statistically calculate that in the one of the next steps.
```{r echo=FALSE}
# Summary statistics
basicStats(MSFT_xts.retDaily)
```

Let's now calculate `average` and `annualized` daily `return` and `volatility`.

Expected annual return is `42.2%` while expected annual volatility is `32.4%`.
```{r echo=FALSE}
# Volatility

avg_return <- mean(MSFT_xts.retDaily)
vol <- stdev(MSFT_xts.retDaily)

print(c("Average daily return:", round(avg_return, 4)))
print(c("Annualized return:", round(avg_return * 252, 3)))
print(c("Volatility:", round(vol, 3)))
print(c("Annualized volatility:", round(vol * sqrt(252), 3)))


```

#### Zero mean test

We calculated that the expected daily return is `0.18%`. Let's do zero mean test
(null hypothesis is that returns are zero on average). We got that `p-value` for 
the test is `0.0335` which is lower than the `significance level of 0.5` (for
rejecting the null hypothesis). Because of the we reject the null hypothesis and
accept the alternative, that the returns on average are not zero.

Previously we already calculated average daily return, which is not zero. 
But because the return is not so large number, we got the p-value for the test to be equal
to 0.033, slightly less than 0.5. That means that if the daily return was smaller than 0.18%, 
it wouldn't be statistically significant and we would probably accept the null hypothesis. 
In that case we could say with statistical certainty that average daily return is 0.
```{r echo=FALSE}
# Testing zero mean return
t.test(MSFT_xts.retDaily) 
```


#### Central moments

Third central moment (`skewness`) is `-0.093`. As skewness is negative and small,
returns are slightly skewed to the left. We'll also perform skewness test to make
sure that it's statistically significant.

Obtained p-value for the test is less than significance level, so we reject the null
hypothesis (H0 is that it's normal distribution, meaning that skewness is 0).
```{r echo=FALSE}
SK = skewness(MSFT_xts.retDaily)  # Skewness
T = length(MSFT_xts.retDaily)     # Sample size
t3 = SK/sqrt(6/T)                 # Skewness test
pval_S = 2*(1-pnorm(abs(t3)))     # p-value

print(c("Skewness:", SK))
print(c("P-value:", pval_S))
```


Fourth central moment (`kurtosis`) is `10.13`. In this case we're calculating
`excess kurtosis`, with the reference value is 0, the same as for the skewness.
As kurtosis is positive and large number, returns distribution is `leptokurtic`.
That means that distribution has `fatter tails`, indicating potential `tail risk`.
Returns that are more than three standard deviations are considered to instantiate
tail risk.

Similarly as for the skewness, we'll also perform kurtosis test. 
Obtained p-value for the test is less than significance level so we reject the null
hypothesis (H0 is that it's normal distribution, meaning that kurtosis is 0).
```{r echo=FALSE}
K = kurtosis(MSFT_xts.retDaily)  # Kurtosis
t4 = K/sqrt(24/T)                # Kurtosis test
pval_K = 2*(1-pnorm(abs(t4)))     # Compute p-value

print(c("Kurtosis:", K))
print(c("P-value:", pval_K))
```

#### Distribution of returns

For simplification purposes, especially in academia, it is often assumed that 
returns follow `normal distribution`. As we previously seen from calculating 
kurtosis, assuming normal distribution could be dangerous because it ignores 
`tail risks` for rare and extreme events.

`Jarque-Bera normality test` is statistical test which checks if distribution
follows normal distribution. For MSFT returns we got that `p-value` is really small
(`2.2e-16`), so we reject the null hypothesis. This test takes both deviation in
the skewness and in the kurtosis. 

```{r echo=FALSE}
normalTest(MSFT_xts.retDaily,method='jb')
```

To have a better felling of all the metrics we calculated previously we can plot
the `histogram of MSFT daily returns`.

Just by looking at the histogram, if we remove outliers, we could say that returns
follow normal distribution. As skewness is nearly equal to zero, histogram looks
quite symmetric.

```{r echo=FALSE}
hist(MSFT_xts.retDaily, nclass=35)
```

In order to better compare distribution of MSFT daily returns to normal distribution
we will estimate PDF and plot it together with normal distribution.

What we previously calculated from Jarque-Bera normality test, we can clearly
see that MSFT returns are deviating from normal distribution.
```{r echo=FALSE}
# Obtain density estimate
d1 = density(MSFT_xts.retDaily)

# Create a sequence of x with increment 0.001.
x = seq(-.4,.4,.001)

# Plot the density
plot(d1$x, d1$y, xlab='rtn', ylab='density', type='l') 

# Create artificial data from Normal distribution
# with mean and variance equal to sample values of MSFT returns
y1 = dnorm(x, mean(MSFT_xts.retDaily), stdev(MSFT_xts.retDaily)) 

# Add density of normal r.v.
lines(x, y1, lty=2)
legend("topright", c("MSFT (full line)", "Normal (dash line)"))
```


#### Comparing to S&P500

In previous chapters we examined returns on MSFT in the great details and 
performing various statistical tests.

Now let's compare the MSFT performance compared to the S&P500 index. Here is the
S&P500 data:

```{r echo=FALSE}
# Importing S&P500 data
SNP <- read.csv(file = "../data/S&p500.csv", row.names = 1, header = TRUE)

# Converting price series to xts objects
# (in order to work with quantmod library)
SNP_xts <- xts(SNP[, 1:5], order.by=as.POSIXct(SNP$date))
SNP_xts.retDaily <- periodReturn(SNP_xts, period = "daily")

# Printing data
tail(SNP[,1:5])

```

We can plot both the MSFT and S&P500 price movement. If you recall our analysis, at 
the beginning, we notices the upward trend in MSFT stock until Q2 of 2020, where
the stock price corrected and from H2 of 2020 it continued with the upward trend.
`The same thing we notice for the S&P500`.

Just by looking at the plot it seems that the MSFT and S&P500 are `higly correlated`,
but in the following steps we'll perform formal statistical test to see what it
the level of the linear dependence.
```{r echo=FALSE}
# Panels
plot.zoo(cbind(MSFT_xts$MSFT.Close, SNP_xts$GSPC.Close), main = "MSFT and S&P500", xlab = "Date", col="blue")
```

#### Regression

In order to find out the level of `liner dependence` between MSFT S&P500 return,
we'll do a `simple liner regression`.

We can see that on average MSFT (`0.175%`) has a higher daily return compared to
the S&P500 (`0.099%`).

As we previously seen just from looking at the plot, it seemed that there was
a `high correlation` between them. Now we calculated `correlation` (standardized
covariance) which is extremely high with value of `0.85`.

```{r echo=FALSE}
rt = cbind(MSFT_xts.retDaily, SNP_xts.retDaily)

# Calculating daily returns
dr = apply(rt, 2, mean)
print("Average daily returns:")
print(c("MSFT (%)", round(100 * dr[[1]], 4)))
print(c("S&P500 (%)", round(100 * dr[[2]], 4)))

# Calculating correlation
correlation = cor(MSFT_xts.retDaily, SNP_xts.retDaily)
print(c("Returns correlation:", correlation))
```

Now let's do the `simple linear regression` of MSFT and S&P500 returns.
We want to find out if there is a `positive/negative association between` MSFT
and S&P500 returns and perform a formal statistical test.

From the regression summary, we got the value for the intercept and for the 
coefficient. The model fitted returns and the value for the coefficient is 
`1.144`. T-test is automatically performed to see if the coefficient has a
statistical significance (null hypothesis is that coefficient is zero).
For the p-value we got the value of `2e-16`, so we could reject the null hypothesis
with a high confidence. 

On the `scatter plot` of the returns which contains the `regression line`  as well,
once again we can clearly see the high level of `positive association` between 
MSFT and S&P500 returns.

```{r echo=FALSE}
# Fitting the linear regression
mreg = lm(MSFT_xts.retDaily ~ SNP_xts.retDaily) 
summary(mreg)

# Scatter plot of return and the regression line
two_returns = cbind(as.numeric(MSFT_xts.retDaily), as.numeric(SNP_xts.retDaily))
plot(two_returns[,1], two_returns[,2], xlab='MSFT', ylab='S&P500', cex=0.8)
abline(0.008, 0.807)
```


#### Conditional PDF and CDF estimation

In this section we'll use estimator from `np library` (Nonparametric Kernel 
Smoothing library) to estimate MFST returns `PDF` and `CDF` conditional on the 
S&P500 returns. We want to find out where is the mass of the density and 
distribution.

As we already knew from previous steps, returns are highly correlated and there
is a positive association between them. That's the reason why the most of the 
mass of the density is located across diagonal, so if S&P500 return is positive
there is a high likelihood that MSFT return will be positive as well. And all
of the peaks are also diagonally distributed.

```{r echo=FALSE, results = FALSE}
# Estimating conditional PDF

# Find the optimal bandwidths
bw <- npcdensbw(xdat=two_returns[,2], ydat=two_returns[,1], tol=.1, ftol=.1)

# Get the model estimate
fhat <- npcdens(bws=bw) 

# Plotting the density
plot(fhat, view = "fixed", main = "", xlab='S&P500',ylab='MSFT', theta = 300,phi = 50)

```

Now we estimate and plot conditional CDF. Here we observe all the similar things
we previously described for the conditional PDF.
```{r echo=FALSE, results = FALSE}
# Estimating conditional CDF

# Find the optimal bandwidths
bwdist <-npcdistbw(xdat=two_returns[,2], ydat=two_returns[,1], tol=.1, ftol=.1)

# Get the model estimate
Fhat <- npcdist(bws=bwdist)

# Plotting the density
plot(Fhat, view = "fixed", main = "", xlab='S&P500',ylab='MSFT', theta = 120, phi = 30)
```
