---
title: "Forecasting multivariate time series models (regression models)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE, warning=FALSE}
# Importing libraries
suppressMessages(library(quantmod))
suppressMessages(library(forecast))
suppressMessages(library(tseries))
```

```{r echo=FALSE}
training_dataset_range <- '2019::2020'
test_dataset_range <- '2021::'
```

## Introduction

The goal of this phase is to produce the best `multivariate regression model` 
for forecasting the return on our stock of choice - Microsoft. For that we will
use family of Linear regression models to find the best performing model.


The dependent variable in our regression model will be daily returns of Microsoft.
The chosen explanatory (independent) variables are also stocks (potential competitors) and stock market stock indexes.

Potential regressors in our regression models are:

- Apple (AAPL)
- Google (GOOG)
- IBM (IBM)
- 3M (MMM)
- S&P500 (^GSPC)
- Nasdaq (^IXIC)


## Splitting the dataset ("in-sample and" "out-of-sample")

The dataset splitting for dependent variable (Microsoft daily returns) has been done in the previous phase.

The training data set will contain daily return data from 2019. and 2020. and the test data will only contain first six months of 2021.


```{r, echo=FALSE}
MSFT <- read.csv(file = "../data/MSFT.csv", row.names = 1, header = TRUE)
MSFT_xts <- xts(MSFT[, 1:5], order.by=as.POSIXct(MSFT$date))
MSFT_xts.retDaily <- periodReturn(MSFT_xts, period = "daily")

MSFT_daily_ret_original <- ts(as.numeric(MSFT_xts.retDaily))
MSFT_daily_ret_training <- ts(as.numeric(MSFT_xts.retDaily[training_dataset_range]), frequency = 252, start=c(2019, 1))
MSFT_daily_ret_test <- ts(as.numeric(MSFT_xts.retDaily[test_dataset_range]), frequency = 252, start=c(2021, 1))

```

In order to split the dataset for potential regressors, we first need to check the stationarity properties of these time series, which is described in the next section.


## Stationarity property of explanatory variables

In this section, we will check the stationarity property of each time series.
That means, it needs to be determined that the time series is constant in mean and variance are constant and not dependent on time.

We will look at couple of methods for checking stationarity:

- Autocorrelation Function (ACF) - Identifying if correlation at different time lags goes to 0
- Augmented Dickey–Fuller (ADF) t-statistic test for unit root
- Kwiatkowski-Phillips-Schmidt-Shin (KPSS) for level or trend stationarity

### Apple (AAPL)
```{r, echo=FALSE}
AAPL <- read.csv(file = "../data/AAPL.csv", row.names = 1, header = TRUE)
AAPL_xts <- xts(AAPL[, 1:5], order.by=as.POSIXct(AAPL$date))

chartSeries(AAPL_xts$AAPL.Close, theme = "white", up.col="blue", name = "AAPL - Closing prices")

# Converting to ts object
AAPL_prices <- ts(as.numeric(AAPL_xts$AAPL.Close))
```

```{r, warning=FALSE, echo=FALSE}
# ACF
acf(AAPL_prices, lag.max = length(AAPL_prices))

# ADF test - closing price
adf.test(AAPL_prices)

# KPSS test - closing price
kpss.test(AAPL_prices, null="Trend")
```

```{r, echo=FALSE}
AAPL_xts.retDaily <- periodReturn(AAPL_xts, period = "daily")
chartSeries(AAPL_xts.retDaily, theme = "white", up.col="blue", name = "AAPL - Daily returns")

AAPL_retDaily <-ts(as.numeric(AAPL_xts.retDaily))
```

```{r, warning=FALSE, echo=FALSE}
# ACF
acf(AAPL_retDaily, lag.max = length(AAPL_retDaily))

# ADF test - closing price
adf.test(AAPL_retDaily)

# KPSS test - closing price
kpss.test(AAPL_retDaily, null="Trend")

```
``` {r, echo=FALSE}
# Splitting the dataset
AAPL_daily_ret_training <- ts(as.numeric(AAPL_xts.retDaily[training_dataset_range]), frequency = 252, start=c(2019, 1))
AAPL_daily_ret_test <- ts(as.numeric(AAPL_xts.retDaily[test_dataset_range]), frequency = 252, start=c(2021, 1))
```